---
title: "Analysis of Adaptive Immune Receptor Repertoire Sequencing data with LymphoSeq2"
author: 
 - "Elena Wu"
 - "Shashidhar Ravishankar"
 - "David Coffey"
date: "`r Sys.Date()`"
output: 
    html_document:
        toc: true
        toc_float: true
        theme: cosmo
        highlight: default
        df_print: paged
vignette: >
  %\VignetteIndexEntry{LymphoSeq2}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Adaptive Immune Receptor Repertoire Sequencing (AIRR-seq) provides a unique opportunity to interrogate the adaptive
immune repertoire under various clinical conditions. The utility offered by this technology has quickly garnered interest
from a community of clinicians and researchers investigating the immunological landscapes of a large spectrum of health
and disease states. LymphoSeq2 is a toolkit that allows users to import, manipulate and visualize AIRR-Seq data from various 
AIRR-Seq assays such as Adaptive ImmunoSEQ and BGI-IRSeq, with support for 10X VDJ sequencing coming soon. The platform 
also supports the importing of AIRR-seq data processed using the MiXCR pipeline. The vignette highlights some of the key features of LymphoSeq2.

# Importing data

The function `readImmunoSeq` imports AIRR-seq receptor files from Adaptive ImmunoSEQ assay as well well as BGI-IRSeq assay. The sequences can be (.tsv) files processed using one of the three following platforms: Adaptive Biotechnologies ImmunoSEQ analyzer, BGI IR-SEQ iMonitor platform, and the MiXCR pipeline for AIRR-seq data analysis. The function has the ability to identify file type based on the headers provided in the (.tsv) file, accordingly the data is transformed into a format that is compatible AIRR-Community guidelines  (https://github.com/airr-community/airr-standards). 

To explore the features of LymphoSeq, this package includes 2 example data sets. The first is a data set of T cell receptor beta (TCRB) sequencing from 10 blood samples acquired serially from a single patient who underwent a bone marrow transplant (Kanakry, C.G., et al. JCI Insight 2016;1(5):pii: e86252). The second, is a data set of B cell receptor immunoglobulin heavy (IGH) chain sequencing from Burkitt lymphoma tumor biopsies acquired from 10 different individuals (Lombardo, K.A., et al. Blood Advances 2017 1:535-544). To improve performance, both data sets contain only the top 1,000 most frequent sequences. The complete data sets are publicly available through Adapatives’ immuneACCESS portal. As shown in the example below, you can specify the path to the example data sets using the command 

```{r data}
system.file("extdata", "TCRB_sequencing", package = "LymphoSeq2") #For the TCRB files
system.file("extdata", "IGH_sequencing", package = "LymphoSeq2") #For the IGH files.
```

`readImmunoSeq` can take as input, a single file name, a list of files or the path to a directory containing AIRR-seq data. The columns are renamed to follow AIRR-community guidelines based on the input file type. The function returns a tibble with individual file names set as the `repertoire_id`. The CDR3 nucleotide and amino acid sequences are denoted by the `junction` and `junction_aa` fields respectively. The counts of the CDR3 sequeneces observed, and their frequency in each individual repertoire is denoted by the `duplicate_count` and `duplicate_frequency` field respectively.

```{r readImmuno}
library(LymphoSeq2)
library(tidyverse)
study_files <- system.file("extdata", "TCRB_sequencing", package = "LymphoSeq2") 
study_table <- LymphoSeq2::readImmunoSeq(path = study_files, threads = 1)
```

Looking at the `study_table` we see a tibble with 12 columns and 9334 rows

```{r showStudyTable}
study_table
```
Since the study table is a tibble, we can use tidyverse syntax to extract a list of sample names

```{r getSamplesNames}
study_table %>% 
dplyr::pull(repertoire_id) %>% 
unique()
```


# Session info

```{r sessionInfo}
sessionInfo()
```
